---
title: "portfolio-piece-1"
author: "Marcus Minko"
date: "1/28/2022"
output: html_document
---
#The goal for this project is too plot and analyze correlations between types of crime and other variables
#data: United States Crime Data

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

### Looking at the dataset USArrests
```{r data-glimpse}
glimpse(USArrests)
```

### Descriptive Statistics
``` {r data-summary}
summary(USArrests)
```

### Creating new variable: total violent crime
``` {r mutate-total-violent-crime}
USArrests <- USArrests %>% 
    mutate(USArrests, 
           violent_crime_total = Murder + Assault + Rape
           )
```

### Top 10 least violent states

```{r least-violent}
USArrests %>% 
    select(violent_crime_total) %>% 
    arrange(violent_crime_total) %>% 
    head(10)
```

### Top 10 most violent states

```{r most-violent}
USArrests %>% 
   select(violent_crime_total) %>% 
   arrange(desc(violent_crime_total)) %>% 
   head(10)
```

### What's the relationship between violent crime and percentage of urban population

```{r plot-violent-crime-urban-population}
ggplot(data = USArrests) +
     geom_point(
       mapping = aes(
                     x = violent_crime_total, 
                     y = UrbanPop
                     )) +
     labs(
          title = "Total Violent Crime by Percentage Urban Population", 
            x = "Total Violent Crime", 
            y = "Percentage Urban Population"
            )
# Correlation between Total Crime and Urban Population
USArrests %>% 
     summarize(r = cor(violent_crime_total, UrbanPop))
```

A correlation of .28 suggests that total crime and percentage of Urban Population are somewhat related.

### Relationship of percentage urban population to each specific criminal act

```{r crime-by-urban-population}
ggplot(data = USArrests) +
     geom_point(
         mapping = aes(
             x = Murder, 
             y = UrbanPop
         )) +
     labs(
         title = "Murders by Percentage Urban Population", 
         x = "Murders (per 100,000)", 
         y = "Percentage Urban Population"
     )
# correlation between murders and percentage urban population
USArrests %>% 
     summarize(r = cor(Murder, UrbanPop))
# A correlation of only .07 suggests that murder is not related to percentage of urban population
ggplot(data = USArrests) +
     geom_point(
         mapping = aes(
             x = Assault, 
             y = UrbanPop
         )) +
     labs(
         title = "Assaults by Percentage Urban Population", 
         x = "Assaults (per 100,000)", 
         y = "Percentage Urban Population"
     )
# Correlation between assaults and percentage urban population
USArrests %>% 
     summarize(r = cor(Assault, UrbanPop))
# A correlation of .26 suggest that there is somewhat of an association between urban population and assaults
ggplot(data = USArrests) +
     geom_point(
         mapping = aes(
             x = Rape, 
             y = UrbanPop
         )) +
     labs(
         title = "Rape Cases by Percentage Urban Population", 
         x = "Rape Cases (per 100,000)", 
         y = "Percentage Urban Population"
     )
# Correlation between assaults and percentage urban population
USArrests %>% 
     summarize(r = cor(Rape, UrbanPop))
# A correlation of .41 suggest that there is pretty good evidence that Rape is positively correlated with the percentage of urban population in a state
```

In Summary, higher murder rates does not appear to be related to a greater percentage of a states population being urban, whereas rape seems to be pretty strongly related to urban population percentage in a positive direction.